|

### 2021-03-24 paper/v2_taylor_francis/response_reviewers.md

> 2. The results for network speed for the different regions are not very different with 1-2 out of 10Mbps difference and probably can not reflect the social disparities. Not sure if this is the facts in U.K. If so, then it's not a problem?

Firstly, mean speeds are not likely to be that different between regions of the UK
(e.g. North or South England). The greater differences are between urban, built-up
areas, and more rural areas within regions, although there are some local initiatives
to address this in some areas. However, we assume that your comment refers to Table 1
and the differences in mean network speed between clusters? Whilst these differences
are not large, the range for upload speeds is smaller than for download speeds,
and a difference of 1-2Mbps slower may well be noticeable at the wrong moment.
Importantly, we only consider the test which took place during working hours, so
when speed drops during, let's say, a work Zoom call or when uploading large files,
it will be noticeable. This is why we created the clusters by how the time series
of speeds changes by day of the week and hour of the day, rather than focusing our
analysis on mean speeds.

> 3. The relevance to pandemic is not very straightforward. This may be improved by adding some studies, e.g., how the clusters impact the effectiveness of work from home through survey. Or relates to the employment rate.

Regrettably, we did not have the opportunity to run a survey. 
As we say in the introduction, prior to the pandemic, the demand for telecommuting 
would not have put a strain on internet connection speeds. We have clarified this 
with the addition of the following sentence:
"High levels of demand are one of the main causes of reduced reliability and slower connection speeds, as network bandwidth becomes congested".
We also have added a variable to the auxiliary regression tracking the share of
the population that was put on furlough. This was the scheme in the UK which, at
least initially, prevented much change in the employment rate, but did show who
could work and who could not at the height of the first wave of the pandemic.
Finally, as we note at the end of the methods and data section, an auxiliary 
regression was the preferred methodology because:
"Such individual data could be found though surveys, but these offer less
detailed information about the experience of internet resilience due to enforced
demand, which is the main contribution of this paper."

@@

We chose the time range to reflect the period of first lockdown in the UK, as we
say in the first paragraph of 3.2:
"We are particularly interested in upload speeds and the frequency of speed tests
over the period from March to May $2020$, as government statements indicate this
encompasses the period when UK workers were first told to work from home if at all
possible [@GovUK2020]".

We have also added the following text:
"Schools and various retail, leisure and hospitality businesses were closed from late
March, and restrictions were gradually eased from late May".

> 5. Would suggest a review of the spatiotemporal challenges of the covid-19.

We have now added the following short discussion in the 'Digital divides and 
economic resilience' section.

"Such digital intersections and efficiencies have also changed 
transportation, connectivity and the structure of cities during the pandemic.
We now have enough data to confirm the drastic alteration observed in the space-time 
geography of cities around the world in the last year [e.g. @google2020; @shibayama2021impact]. 
Yet, although there is a broad agreement that these changes 
during the pandemic have played a pivotal role in stopping the spread of the virus 
[@jia2020population; @yang2020taking; @mu2020interplay], the extent to which the 
increased levels of working from home, the consequent decrease in commuting flows, 
and the altered structure of cities will remain post-pandemic is the subject
of considerable debate in the literature."

@@

> 1. The paper used k-means, a popular unsupervised classification in this study. The optimal number of clusters was determined as 13 based on cluster validation indices (CVIs). While this number could be well justified, I feel it is too high to properly interpret the results. The interpretation and knowledge discovery from clustering results with a high k tend to be tedious. Being affected by the high k, the Results section appears to be very difficult to read and follow. As there are many different CVIs, I feel a lower k will benefit the interpretation of results.

Please see our response to the next point.

> 2. Another major comment is on the multinomial logit regression (MLR). This is partly related to my first comment on a larger k. While there is no hard limit on the number of categories in MLR, the interpretation of results will be difficult with a high k as well. In addition, if the dependence among independent variables (collinearity) occurs, the estimation of the multinomial logit model parameters becomes inaccurate. According to Table 2, it is very likely some of the variables could present collinearity (e.g., distance to nearest metro and tech jobs).

Following your suggestion, we have now opted to present the cluster analysis and
the subsequent multinomial regression using $k = 9$. We indeed believe that the
results are more clear now. we have added the following justification in the Methods
and Data section:

@@

> 3. Although Fig. 1 and 2 exhibit different speed tests patterns in 2019 and 2020, the data mining outcome lacks a baseline (prior to 2020) to compare with. It is very likely that the clusters in 2019 would be different from that in 2020. It will make the results more robust. Of course, I will not push this given the page limit. But some insights may be helpful in the Discussion (however, a Discussion section is missing).

ET: after we finalise k, we can rerun k-means for 2019 and decide - could also show mean speeds on a map for 2019 - this would be part of the response to reviewer 1 q2 and q6.
**Let's discuss this**

> 4. No discussion was included in this manuscript. I expect some insights into various important issues, such as problem causes, any broadband policies and socioeconomic solutions, equitable development, and research uncertainties.

We have tightened up the results section, and begun this discussion in the last
paragraph of the results section and then expanded the conclusion section to include
a more in-depth discussion of the results and how they link to some of the issues you 
suggest at the start. We mention some uncertainties or avenues for future research,
but also see our response to Reviewer 1, point 1 for an explanation of biases.

### 2021-03-25 paper/v2_taylor_francis/response_reviewers.md

"We initially run the algorithm for $k \in \mathbb{N} \bigcap [5,15]$, calculated
the cluster validity indices (CVIs) and then run the
subsequent multinomial regression -- see the end of this section for more details
for the different *k*.
Following @sardatime, to identify the optimal *k* we used the majority vote for
the following CVIs: Silhouette (max), Score function (max), Calinski-Harabasz (max),
Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).
Nevertheless, we opted against using the optimal $k = 13$ solution as it was too
large to allow for communicable LAD clusters. Instead we opted for a smaller $k = 9$,
which led to a rather similar spatial pattern and, importantly, to a much higher
R-squared in the subsequent explanatory regression ($0.44$ instead of $0.34$)."

Regarding multicolinearity, the highest correlations were observed between $NVQ4+$
and $professional jobs$ ($0.78$) and $NVQ4+$ and earnings ($0.69$). Hence, we decided
to exclude $NVQ4+$ from the regression.

> 3. Although Fig. 1 and 2 exhibit different speed tests patterns in 2019 and 2020, the data mining outcome lacks a baseline (prior to 2020) to compare with. It is very likely that the clusters in 2019 would be different from that in 2020. It will make the results more robust. Of course, I will not push this given the page limit. But some insights may be helpful in the Discussion (however, a Discussion section is missing).

This is useful. We have not added a new Appendix 2, which plots the clusters for 
the same period in 2019. The picture is very different, which is an indication of 
the changes in internet usage during the pandemic.

"It is worth noting here that clusters of LAD based on the upload speeds are very 
different to the ones occur when we apply the same methods to upload speed data
for the same time period in 2019 -- see Appendix 2. This drastic 
difference between the LAD upload speed profiles before and during the pandemic is
an indication of the changes in internet usage that took place during the pandemic."

### 2021-03-25 paper/v2_taylor_francis/broadband.speed.covid.Rmd
new appendix with 2019 clusters, text for discussion, new images of higher res, ref and footnote for openair added

<!-- data transformation -->
The frequency of speed tests was important in identifying the temporal profile which
would give us most insight into experienced  internet service and resilience over
units of time.
Whilst there is an overall trend of increased testing
from March to April and then a slight reduction from April to May,
this trend masks substantial variation by not only the day of the week,
but also time of day, as can be seen in Figure \ref{test2020}.
Thus, a daily aggregation of upload speeds would mask the variation in experienced
service over the course of each weekday.
Furthermore, the importance of this variation is highlighted by a comparison
with the same period in $2019$, as in Figure \ref{test2019}, when the volume of
testing and thus of experienced internet service quality peaked in the evening,
presumably in response to demand for leisure activities and download speeds.
In contrast, the majority of the increase in testing in $2020$ is during the working
day, creating a new morning peak in Figure \ref{test2020}.
Therefore, we include a measure of hourly variation in our temporal profiles to
reflect the change in users' perception of the workday reliability of internet
services.

@@

@@

First, we control for the number of speed tests run per cluster inhabitant
between $9:00$-$10:59$ as well as the share of fast Virgin Media internet
connections. Regarding the former, we expect people in LADs with more unreliable
connections to test their internet speeds more often, and the results show that 
those in cluster $6$ are by far the most concerned about reliability at that time 
of day. Meanwhile, those in clusters $2$, $3$, and $9$ benefit from a higher 
proportion of Virgin connections, which is an indication that people in these 
clusters are more likely to live in urban areas, with more profitable broadband 
markets, and a better choice of broadband services. 
We also employ distance from London and from the nearest metropolitan area
(including London) as two variables depicting peripherality and urban structure.
However, whilst significant, the pattern of coefficients of these variables is 
inconclusive in confirmed first level digital divides. 
Even though London was one of the ten largest metropolitan areas in 
England, which, along with Glasgow and Cardiff, were identified to calculate the 
variable estimating the impact of distance from the centre of a metropolitan area, 
the coefficients for the two variables are mostly small and the signs for some are 
opposite.

@@

<!-- WFH broader consequences-->
Therefore, home-based digital infrastructure which considers upload speeds and 
working day reliability as well as availability are likely to be particularly 
important in a future where telecommuting might be a more common means of accessing 
work and broadband services must be fit for purpose. 
Although the long-term effects of such drastic changes in telecommuting and 
attitudes towards working from home are difficult to predict, the reliability of 
home broadband services deserve more consideration than in the past. 
This would represent a switch from previous broadband policies in the UK which 
tended to be aimed at supporting small and medium enterprises on industrial estates 
and business parks []. Such policies also made some assumptions about the quality of 
service in residential markets that could be assured without government intervention 
- an assumption which our analysis suggest is not entirely accurate.

@@

<!-- \pagebreak -->

@@

@@

>>>>

<!-- data transformation -->
The frequency of speed tests was important in identifying the temporal profile which
would give us most insight into experienced  internet service and resilience over
units of time.
Whilst there is an overall trend of increased testing
from March to April and then a slight reduction from April to May,
this trend masks substantial variation by not only the day of the week,
but also time of day, as can be seen in Figure \ref{test2020}
^[Figures \ref{test2020} and \ref{test2019} were created with the `R` package `openair` [@openair].].
Thus, a daily aggregation of upload speeds would mask the variation in experienced
service over the course of each weekday.
Furthermore, the importance of this variation is highlighted by a comparison
with the same period in $2019$, as in Figure \ref{test2019}, when the volume of
testing and thus of experienced internet service quality peaked in the evening,
presumably in response to demand for leisure activities and download speeds.
In contrast, the majority of the increase in testing in $2020$ is during the working
day, creating a new morning peak in Figure \ref{test2020}.
Therefore, we include a measure of hourly variation in our temporal profiles to
reflect the change in users' perception of the workday reliability of internet
services.

@@

It is worth noting here that clusters of LAD based on the upload speeds are very 
different to the ones occur when we apply the same methods to upload speed data
for the same time period in 2019 -- see Appendix [2](#appendix2). This drastic 
difference between the LAD upload speed profiles before and during the pandemic is
an indication of the changes in internet usage that took place during the pandemic.

@@

First, we control for the number of speed tests run per cluster inhabitant
between $9:00$-$10:59$ as well as the share of fast Virgin Media internet
connections
^[See Appendix [3](#appendix3) for the descriptive statistics.].
Regarding the former, we expect people in LADs with more unreliable
connections to test their internet speeds more often, and the results show that 
those in cluster $6$ are by far the most concerned about reliability at that time 
of day. Meanwhile, those in clusters $2$, $3$, and $9$ benefit from a higher 
proportion of Virgin connections, which is an indication that people in these 
clusters are more likely to live in urban areas, with more profitable broadband 
markets, and a better choice of broadband services. 
We also employ distance from London and from the nearest metropolitan area
(including London) as two variables depicting peripherality and urban structure.
However, whilst significant, the pattern of coefficients of these variables is 
inconclusive in confirmed first level digital divides. 
Even though London was one of the ten largest metropolitan areas in 
England, which, along with Glasgow and Cardiff, were identified to calculate the 
variable estimating the impact of distance from the centre of a metropolitan area, 
the coefficients for the two variables are mostly small and the signs for some are 
opposite.

@@

<!-- WFH broader consequences-->
Therefore, home-based digital infrastructure which considers upload speeds and 
working day reliability as well as availability are likely to be particularly 
important in a future where telecommuting might be a more common means of accessing 
work and broadband services must be fit for purpose. 
Although the long-term effects of such drastic changes in telecommuting and 
attitudes towards working from home are difficult to predict, the reliability of 
home broadband services deserve more consideration than in the past. 
This would represent a switch from previous demand-side broadband policies in the UK
which tended to be aimed at supporting small and medium enterprises [@HENDERSON2020102024].
SUch policies were based on previous research regarding the productivity effects 
of broadband infrastructure [@DESTEFANO2018110]. However, this stream of research 
tended not to consider residential locations as places that host economic activity.
Such policies also made some assumptions about the quality of 
service in residential markets that could be assured without government intervention 
-- an assumption which our analysis suggest is not entirely accurate.

@@

\pagebreak

@@

```{r echo=FALSE, out.width = "100%",fig.cap="\\label{map.up.2019.clusters}Upload speed clusters for LADs in before the pandemic (2019)"}
img_path <- "figures/map.up.2019.clusters.png"
include_graphics(img_path)
```

\pagebreak

# Appendix 3 {- #appendix3}

@@

\pagebreak

# References
